---
title: "Aalen-Johansen model estimates"
output:
    html_document:
        code_folding: hide
        toc:          true
---

Aalen-Johansen non-parametric estimation is used to estimate the competing risks of death and discharge following hospital admission, and estimate median length of stay in hospital.

```{r setup, include=FALSE}

here::i_am("rmd/aj_model.Rmd")

library(tidyverse)
library(lubridate)
library(patchwork)
library(here)
library(survminer)

source(here("r/aj_functions.R"))

load(here("data/processed_data.RData"))

figure_path <- here("figures")
```

```{r nd_datasets}

# ordered factor data
level_order_thinned <- c("Mar 2020", "May 2020", "Jul 2020", "Sep 2020", "Nov 2020", "Jan 2021", "Mar 2021", "May 2021", "Jul 2021", "Sep 2021")
age_order <- c("0-14", "15-24", "25-44", "45-64", "65-74", "75-84", "85+")
phec_order <- c("London", "East Midlands", "East of England", "North East", "North West", "South East", "South West", "West Midlands", "Yorkshire and Humber")

# generate expanded tibbles
nd_month <- tibble(monthyear = level_order)
nd_month_sex <- expand_grid(nd_month, sex = c("Male", "Female"))
nd_month_age <- expand_grid(nd_month, ageGrp7 = age_order)
nd_month_phe <- expand_grid(nd_month, pheRegion = phec_order)
nd_month_ethn <- expand_grid(nd_month, ethGrp4 = c("White", "Asian", "Black", "Mixed/Other/Unknown"))
nd_month_imd <- expand_grid(nd_month, imd_quintile = c("1", "2", "3", "4", "5"))
nd_month_cci <- expand_grid(nd_month, charlson_index = c("0", "1-2", "3-4", "5+"))
nd_month_load <- expand_grid(nd_month, bed_occupancy_cat = c("0-20%", "20-40%", "40-60%", "60-80%", "80-90%", "90-100%"))
```

## Aalen-Johansen model results

Hospitalised fatality risk (panel A) and median length of stay (panel B) by month of admission and one other covariate.

```{r overall-los}

aj_full_fit(hosp, nd = nd_month, save_name = "aj_month")

aj_full_fit(hosp, nd = nd_month_sex, save_name = "aj_month_sex")

aj_full_fit(hosp, nd = nd_month_age, save_name = "aj_month_age", col = TRUE, flip = TRUE)

aj_full_fit(hosp, nd = nd_month_phe, save_name = "aj_month_region", col = TRUE, flip = TRUE, fig_height = 19)

aj_full_fit(hosp, nd = nd_month_ethn, save_name = "aj_month_ethn")

aj_full_fit(hosp, nd = nd_month_imd, save_name = "aj_month_imd")

aj_full_fit(hosp, nd = nd_month_cci, save_name = "aj_month_cci")

aj_full_fit(hosp, nd = nd_month_load, save_name = "aj_month_load", col = TRUE, flip = TRUE)
```

Aalen-Johansen and Fine-Gray cumulative fatality estimates for first 60 days following hospital admission for selected months.

```{r survival-curve}

fg_data <- hosp %>% mutate(
    eventm = as.numeric(eventm),
    eventm = if_else(is.na(eventm), 0, eventm),
    eventm = factor(eventm, 0:2, labels = c("Censor", "Death", "Discharge"))
)

pdata <- finegray(Surv(fg_data$time1m, fg_data$eventm) ~ ., data = fg_data)
pdata$monthyear <- relevel(pdata$monthyear, ref = "Jun 2020")

# to view survival curves
ajfit_month <- aj_fit(hosp, nd = nd_month)
fg_fit <- coxph(Surv(fgstart, fgstop, fgstatus) ~ monthyear,
    weight = fgwt,
    data = pdata
)

sf <- survfit(fg_fit, newdata = nd_month)

ss <- survminer::surv_summary(sf, data = pdata) %>%
    transmute(
        time,
        val = 1 - surv,
        lower_s = 1 - upper,
        upper_s = 1 - lower,
        state = "Death",
        strata = as.numeric(strata)
    ) %>%
    rename(lower = lower_s, upper = upper_s)

ss <- ss %>%
    mutate(
        monthyear = factor(strata, labels = number_order)
    )

ajfit_month %>%
    mutate(model = "Aalen-Johansen") %>%
    bind_rows(ss %>% mutate(model = "Fine and Gray")) %>%
    filter(state == "Death") %>%
    mutate(monthyear = factor(monthyear, levels = level_order)) %>%
    filter(monthyear %in% level_order_thinned) %>%
    ggplot() +
    aes(x = time, y = val, ymin = lower, ymax = upper, col = model, fill = model) +
    geom_line() +
    geom_ribbon(alpha = 0.3, lty = 0) +
    labs(caption = "Cumulative probabilities shown for first 60 days following hospital admission.") +
    xlab("Days after hospital admission") +
    ylab("Probability of fatality") +
    scale_x_continuous(breaks = c(0, 25, 50), limits = c(0, 60)) +
    theme_bw() +
    theme(legend.title = element_blank(), legend.position = "top") +
    theme(text = element_text(size = 16)) +
    facet_wrap(~monthyear, nrow = 2)

ggsave(here(figure_path, "fg_aj_month_comparison.pdf"), device = cairo_pdf, width = 12, height = 13)
```
