---
title: "Population characteristics"
output:
    html_document:
        code_folding: hide
        toc:          true
---

```{r setup, include=FALSE}

here::i_am("rmd/pop_chars.Rmd")

library(tidyverse)
library(lubridate)
library(scales)
library(ggpubr)
library(here)

load(here("data/processed_data.RData"))

figure_path <- here("figures")
```

## Tables

Characteristics of the study population compared with all people hospital-onset COVID-19 in England and all people with PCR-confirmed community-acquired COVID-19 in England.

```{r baseline}

hosp %>% count()

# censored outcomes
hosp %>%
    filter(is.na(eventm)) %>%
    count()
hosp %>%
    filter(is.na(eventm) & time1m == 90) %>%
    count()
hosp %>%
    filter(is.na(eventm) & time1m < 90) %>%
    count()

hosp %>%
    group_by(ageGrp7) %>%
    summarise(n = n()) %>%
    mutate(pct = round(n / sum(n) * 100, 1))

hosp %>%
    group_by(sexHC) %>%
    summarise(n = n()) %>%
    mutate(pct = round(n / sum(n) * 100, 1))

hosp %>%
    group_by(ethGrp4) %>%
    summarise(n = n()) %>%
    mutate(pct = round(n / sum(n) * 100, 1))

hosp %>%
    group_by(pheRegion) %>%
    summarise(n = n()) %>%
    mutate(pct = round(n / sum(n) * 100, 1))

hosp %>%
    group_by(imd_quintile) %>%
    summarise(n = n()) %>%
    mutate(pct = round(n / sum(n) * 100, 1))

hosp %>%
    group_by(monthyear) %>%
    summarise(n = n()) %>%
    mutate(pct = round(n / sum(n) * 100, 1))

hosp %>%
    filter(monthyear %in% c("Jan 2021", "Feb 2021", "Mar 2021", "Apr 2021", "May 2021", "Jun 2021", "Jul 2021", "Aug 2021", "Sep 2021", "Oct 2021", "Nov 2021", "Dec 2021")) %>%
    group_by(vaccine) %>%
    summarise(n = n()) %>%
    mutate(pct = round(n / sum(n) * 100, 1))

hosp %>%
    filter(!is.na(bed_occupancy_cat)) %>%
    group_by(bed_occupancy_cat) %>%
    summarise(n = n()) %>%
    mutate(pct = round(n / sum(n) * 100, 1))

hosp %>%
    filter(!is.na(charlson_index)) %>%
    group_by(charlson_index) %>%
    summarise(n = n()) %>%
    mutate(pct = round(n / sum(n) * 100, 1))

hosp %>%
    mutate(linkset = if_else(linkset == "CoV:SUS:ECDS", "CoV:ECDS", linkset)) %>%
    group_by(linkset) %>%
    summarise(n = n()) %>%
    mutate(pct = round(n / sum(n) * 100, 1))

hosp %>%
    group_by(eventm) %>%
    summarise(n = n()) %>%
    mutate(pct = round(n / sum(n) * 100, 1))

hosp %>%
    group_by(eventm) %>%
    summarise(
        med = median(time1m),
        lq = quantile(time1m, 0.25),
        uq = quantile(time1m, 0.75)
    )

# hosp by event
hosp %>%
    group_by(ageGrp7, eventm) %>%
    summarise(n = n()) %>%
    mutate(
        pct = round(n / sum(n) * 100, 1),
        val = paste0(n, " (", pct, "%)")
    ) %>%
    select(eventm, ageGrp7, val) %>%
    pivot_wider(names_from = eventm, values_from = val)

hosp %>%
    group_by(sexHC, eventm) %>%
    summarise(n = n()) %>%
    mutate(
        pct = round(n / sum(n) * 100, 1),
        val = paste0(n, " (", pct, "%)")
    ) %>%
    select(eventm, sexHC, val) %>%
    pivot_wider(names_from = eventm, values_from = val)

hosp %>%
    group_by(ethGrp4, eventm) %>%
    summarise(n = n()) %>%
    mutate(
        pct = round(n / sum(n) * 100, 1),
        val = paste0(n, " (", pct, "%)")
    ) %>%
    select(eventm, ethGrp4, val) %>%
    pivot_wider(names_from = eventm, values_from = val)

hosp %>%
    group_by(pheRegion, eventm) %>%
    summarise(n = n()) %>%
    mutate(
        pct = round(n / sum(n) * 100, 1),
        val = paste0(n, " (", pct, "%)")
    ) %>%
    select(eventm, pheRegion, val) %>%
    pivot_wider(names_from = eventm, values_from = val)

hosp %>%
    group_by(imd_quintile, eventm) %>%
    summarise(n = n()) %>%
    mutate(
        pct = round(n / sum(n) * 100, 1),
        val = paste0(n, " (", pct, "%)")
    ) %>%
    select(eventm, imd_quintile, val) %>%
    pivot_wider(names_from = eventm, values_from = val)

hosp %>%
    group_by(monthyear, eventm) %>%
    summarise(n = n()) %>%
    mutate(
        pct = round(n / sum(n) * 100, 1),
        val = paste0(n, " (", pct, "%)")
    ) %>%
    select(eventm, monthyear, val) %>%
    pivot_wider(names_from = eventm, values_from = val)

hosp %>%
    group_by(bed_occupancy_cat, eventm) %>%
    summarise(n = n()) %>%
    mutate(
        pct = round(n / sum(n) * 100, 1),
        val = paste0(n, " (", pct, "%)")
    ) %>%
    select(eventm, bed_occupancy_cat, val) %>%
    pivot_wider(names_from = eventm, values_from = val)

hosp %>%
    filter(
        monthyear %in% c("Jan 2021", "Feb 2021", "Mar 2021", "Apr 2021", "May 2021", "Jun 2021", "Jul 2021", "Aug 2021", "Sep 2021", "Oct 2021", "Nov 2021", "Dec 2021"),
        vaccine != ">14 days post third dose"
    ) %>%
    group_by(vaccine, eventm) %>%
    summarise(n = n()) %>%
    mutate(
        pct = round(n / sum(n) * 100, 1),
        val = paste0(n, " (", pct, "%)")
    ) %>%
    select(eventm, vaccine, val) %>%
    pivot_wider(names_from = eventm, values_from = val)

hosp %>%
    group_by(charlson_index, eventm) %>%
    summarise(n = n()) %>%
    mutate(
        pct = round(n / sum(n) * 100, 1),
        val = paste0(n, " (", pct, "%)")
    ) %>%
    select(eventm, charlson_index, val) %>%
    pivot_wider(names_from = eventm, values_from = val)

hosp %>%
    mutate(linkset = if_else(linkset == "CoV:SUS:ECDS", "CoV:ECDS", linkset)) %>%
    group_by(linkset, eventm) %>%
    summarise(n = n()) %>%
    mutate(
        pct = round(n / sum(n) * 100, 1),
        val = paste0(n, " (", pct, "%)")
    ) %>%
    select(eventm, linkset, val) %>%
    pivot_wider(names_from = eventm, values_from = val)
```

Nosocomial comparison

```{r noso}

# nosocomial infections

load(here("data/nosocomial_data.RData"))

noso %>% count()

noso %>%
    mutate(
        ageGrp7 = cut(age,
            breaks = c(0, 15, 25, 45, 65, 75, 85, Inf),
            right = FALSE,
            include.lowest = TRUE, ordered.result = TRUE
        )
    ) %>%
    group_by(ageGrp7) %>%
    summarise(n = n()) %>%
    mutate(pct = round(n / sum(n) * 100, 1))

noso %>%
    group_by(sexHC) %>%
    summarise(n = n()) %>%
    mutate(pct = round(n / sum(n) * 100, 1))

noso %>%
    mutate(
        ethnicity_final = factor(ethnicity_final, ordered = FALSE),
        ethGrp4 = fct_collapse(ethnicity_final,
            Asian = c("Any other Asian background", "Bangladeshi (Asian or Asian British)", "Indian (Asian or Asian British)", "Indian (Asian or Asian British)", "Pakistani (Asian or Asian British)"),
            Black = c("African (Black or Black British)", "Any other Black background", "Caribbean (Black or Black British)"),
            `Mixed/Other/Unknown` = c("Any other ethnic group", "Any other Mixed background", "Chinese (other ethnic group)", "White and Asian (Mixed)", "White and Black African (Mixed)", "White and Black Caribbean (Mixed)", "Unknown"),
            White = c("Any other White background", "British (White)", "Irish (White)")
        ),
        ethGrp4 = fct_relevel(ethGrp4, "White", "Asian", "Black", "Mixed/Other/Unknown")
    ) %>%
    group_by(ethGrp4) %>%
    summarise(n = n()) %>%
    mutate(pct = round(n / sum(n) * 100, 1))

noso %>%
    mutate(pheRegion = factor(phec_name, levels = c("London", "East Midlands", "East of England", "North East", "North West", "South East", "South West", "West Midlands", "Yorkshire and Humber"), ordered = FALSE)) %>%
    group_by(pheRegion) %>%
    summarise(n = n()) %>%
    mutate(pct = round(n / sum(n) * 100, 1))

noso %>%
    mutate(imd_quintile = ceiling(imd_decile / 2) %>% factor()) %>%
    group_by(imd_quintile) %>%
    summarise(n = n()) %>%
    mutate(pct = round(n / sum(n) * 100, 1))

noso %>%
    mutate(
        monthyear = paste0(lubridate::month(specimen_date, label = TRUE), " ", year(specimen_date)),
        monthyear = factor(monthyear, levels = level_order)
    ) %>%
    group_by(monthyear) %>%
    summarise(n = n()) %>%
    mutate(pct = round(n / sum(n) * 100, 1))

noso %>%
    group_by(eventm) %>%
    summarise(n = n()) %>%
    mutate(pct = round(n / sum(n) * 100, 1))

noso %>%
    group_by(eventm) %>%
    summarise(
        med = median(time1m),
        lq = quantile(time1m, 0.25),
        uq = quantile(time1m, 0.75)
    )
```

Comparison to all 

```{r comparison}

load(here("data/all_admissions_data.RData"))

all_comparison %>% count()

all_comparison %>%
    group_by(ageGrp7) %>%
    summarise(n = n()) %>%
    mutate(pct = round(n / sum(n) * 100, 1))

all_comparison %>%
    group_by(sex) %>%
    summarise(n = n()) %>%
    mutate(pct = round(n / sum(n) * 100, 1))

all_comparison %>%
    group_by(ethGrp4) %>%
    summarise(n = n()) %>%
    mutate(pct = round(n / sum(n) * 100, 1))

all_comparison %>%
    group_by(pheRegion) %>%
    summarise(n = n()) %>%
    mutate(pct = round(n / sum(n) * 100, 1))

all_comparison %>%
    group_by(imd_quintile) %>%
    summarise(n = n()) %>%
    mutate(pct = round(n / sum(n) * 100, 1))

all_comparison %>%
    group_by(monthyear) %>%
    summarise(n = n()) %>%
    mutate(pct = round(n / sum(n) * 100, 1))
```

## Additional figures

Observed number of individuals hospitalised with COVID-19, by week of admission, March 2020 to September 2021, with annotations for wave 1, wave 2, and wave 3.

```{r hosp-fig}

hosp %>%
    ggplot() +
    aes(week) +
    geom_bar(fill = "darkgreen") +
    geom_bracket(
        xmin = "2020-08", xmax = "2020-28", y.position = 13000,
        label = "Wave 1", label.size = 5, size = 0.7
    ) +
    geom_bracket(
        xmin = "2020-32", xmax = "2021-16", y.position = 13000,
        label = "Wave 2", label.size = 5, hjust = 1, size = 0.7
    ) +
    geom_bracket(
        xmin = "2021-20", xmax = "2021-45", y.position = 13000,
        label = "Wave 3", label.size = 5, size = 0.7
    ) +
    theme_minimal() +
    scale_y_continuous(labels = scales::comma_format()) +
    scale_x_discrete(
        breaks = c("2020-12", "2020-34", "2021-02", "2021-24"),
        labels = c("March\n2020", "August\n2020", "January\n2021", "June\n2021")
    ) +
    labs(y = "", x = "") +
    theme(
        panel.grid.major.x = element_blank(),
        text = element_text(size = 16),
        axis.text.x = element_text(vjust = 7)
    )

ggsave(here(figure_path, "wave_figure.pdf"), width = 14, height = 8)
```

Vaccination status of hospitalised individuals, by month of admission and age group, December 2020 to September 2021.

```{r vaccine-fig, fig.width=15}

vaccine_table <- hosp %>%
    mutate(ageGrp7 = factor(ageGrp7, labels = c("0-15", "15-25", "25-45", "45-65", "65-75", "75-85", "85+"))) %>%
    filter(
        monthyear == "Dec 2020" | grepl("2021", monthyear),
        vaccine != "\u226514 days post third dose"
    ) %>%
    mutate(
        vaccine = fct_drop(vaccine),
        vaccine = factor(vaccine, labels = c("Unvaccinated", "<21 days after first dose", "\u226521 days after first dose", "\u226514 days after second dose"))
    ) %>%
    group_by(monthyear, ageGrp7, vaccine) %>%
    summarise(n = n(), .groups = "drop_last") %>%
    mutate(pct = n / sum(n))

# code for changing one category colour
gg_color_hue <- function(n) {
    hues <- seq(15, 375, length = n + 1)
    hcl(h = hues, l = 65, c = 100)[1:n]
}
cols <- gg_color_hue(4)
cols[1] <- "#F1F1F1"

vaccine_table %>% ggplot() +
    aes(monthyear, n, fill = vaccine) +
    geom_col(position = position_fill()) +
    labs(fill = "", y = "Proportion", x = "") +
    scale_y_continuous(labels = scales::percent, expand = expansion(mult = c(0, 0))) +
    theme_bw(16) +
    theme(
        legend.position = "top",
        axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)
    ) +
    facet_grid(~ageGrp7) +
    scale_fill_manual(values = cols)

ggsave(here(figure_path, "vaccine_month.pdf"), width = 14, height = 8)
```